\chapter{About \rtos{}}
\label{secAbout}

\rtos{} is a small real time operating system (RTOS) for the Arduino
environment. It is simple to use and fits well into the existing Arduino
code environment. It adds the concept of pseudo-parallel execution threads
to your sketches. These threads may be preemptive or non-preemptive.
Deliberate, cooperative switching between the tasks is supported or time
based switching. Events can be sent and received by tasks as a simple kind
of communication.

\rtos{} is provided as a single source code file which is compiled together
with your other code, which now becomes an \rtos{} application. In the most
simple case, if you do not define any task, your application will strongly
resemble a traditional sketch: You implement your \ident{setup} and your
\ident{loop} function; the former will be run once at the beginning and
the latter repeatedly.

\rtos{} on its own can't be compiled, there need to be an application.
\rtos{} is organized as a package which combines the \rtos{} source file
with some sample applications which are the test cases at the same time.
The source code of each sample application is held in a separate folder,
named tc\textless nn\textgreater. Any of these can be selected for
compilation. You may add more folders, holding the source code of your
\rtos{} application. A starting point of your application folder can be a
copy of any of the folders tc\textless nn\textgreater. The compilation
always is the same. Run the makefile, where the name of the folder (which
doesn't need to be tc\textless nn\textgreater) is an option of the
command line. See below for more.

A characteristic of \rtos{} is that the behavior of a task is not
predetermined or fixed at runtime. \rtos{} supports regular,
time-controlled tasks as well as purely event controlled ones. It supports
preemptive tasks and non-preemptive ones. Task scheduling can be done
cooperatively or time based, and round robin like time slices are also
possible. Moreover, many of these modes can be mixed. A task is not per-se
regular, its implementing code decides what happens and this can be
decided context or situation dependent. This flexibility is achieved by
the basic idea of having an event controlled scheduler, where typical RTOS
use cases are supported by providing according events, e.g.
absolute-point-in-time-reached. If the task's code decides to always wait
for the same absolute-point-in-time-reached event, then it becomes a
regular task. However, situation dependent the same task could decide to
wait for an application sent event -- and give up its regular behavior. In
many RTOS implementations the basic characteristic of a task is determined
at compile time, in \rtos{} this is done partly at compile time and partly
at runtime.
